\section{Full user study results}\label{sec:full_user_study_results}
Figures \ref{fig:alltextures_hist_short} and \ref{fig:alltextures_hist_long}
show histograms of the average user accuracy on each texture, averaged over a 
range of exposure times. The histogram bars are ordered from lowest
to highest accuracy, based on the results when using the final model.

Tables \ref{tab:table_concat} and \ref{tab:table_concat_range} show
the average user accuracy on each texture when using the final model.
The results are averaged over exposure times. Similarly, Tables
\ref{tab:table_flowdecode} and \ref{tab:table_flowdecode_range} show the results
when using the baseline.

Tables \ref{tab:table_concat_app} and \ref{tab:table_concat_app_range} show
the average user accuracy on texture appearance groups when using the
final model. The results are averaged over exposure times. Similarly, Tables
\ref{tab:table_flowdecode_app} and \ref{tab:table_flowdecode_app_range} show the results when using the baseline.

Tables \ref{tab:table_concat_dyn} and \ref{tab:table_concat_dyn_range} show
the average user accuracy on texture dynamics groups when using the
final model. The results are averaged over exposure times. Similarly, Tables
\ref{tab:table_flowdecode_dyn} and \ref{tab:table_flowdecode_dyn_range} show the results when using the baseline.

Tables \ref{tab:table_concat_all} and \ref{tab:table_concat_all_range} show
the average user accuracy over all textures when using the
final model. The results are averaged over exposure times. Similarly, Tables
\ref{tab:table_flowdecode_all} and \ref{tab:table_flowdecode_all_range} show the results when using the baseline.