\section{Qualitative results}
Provided in the supplemental material are videos showcasing the qualitative results of the two-stream model, including the experiments mentioned in the main manuscript. The videos are in MP4 format (H.264 codec) and are best viewed in a loop. They are enclosed in the following folders in the supplemental:
\begin{itemize}
	\item \path{target_textures}: This folder contains the 59 dynamic textures used as targets for synthesis.
	\item \path{dynamic_texture_synthesis}: This folder contains synthesized dynamic textures where the appearance and dynamics targets are the same. \highlight{Specifically, it contains the folders \path{comparisons}, \path{using_concatenation_layer}, and \path{using_flow_decode_layer}.}
	\item \path{using_concatenation_layer}: This folder contains synthesized dynamic textures where the concatenation layer was used for computing the Gram matrices on the dynamics stream. These are the results from the final model. \highlight{It contains the folders \path{appearance_stream_only}, \path{dynamics_stream_only}, \path{full_synthesis}, \path{incrementally_synthesized}, and \path{temporally_endless}.}
	\item \path{using_flow_decode_layer}: This folder contains synthesized dynamic textures where the predicted flow output is used for computing the Gram matrices on the dynamics stream. These are the results from the baseline. \highlight{It contains the folder \path{full_synthesis}.}
	\item \path{full_synthesis}: This folder contains regularly-synthesized dynamic textures, \ie, not incrementally-generated, nor temporally-endless, etc.
	\item \path{appearance_stream_only}: This folder contains dynamic textures synthesized using only the appearance stream of the two-stream model. The dynamics stream is not used.
	\item \highlight{\path{dynamics_stream_only}: This folder contains a dynamic texture synthesized using only the dynamics stream of the two-stream model. The appearance stream is not used.}
	\item \path{incrementally_synthesized}: This folder contains dynamic textures synthesized using the incremental process outlined in Sec.\ \ref{sec:incremental_synthesis} in the main manuscript.
	\item \path{temporally_endless}: This folder contains a synthesized dynamic texture (\path{smoke_plume_1}) where there is no discernible temporal seam between the last and first frames. Played as a loop, it appears to be temporally endless, thus, it is presented in animated GIF format.
	\item \path{dynamics_style_transfer}: This folder contains synthesized dynamic textures where the appearance and dynamics targets are different. Also included are videos where the synthesized dynamic texture is ``pasted'' back onto the original image it was cropped from, showing a proof-of-concept of dynamics style transfer as an artistic tool.
	\item \path{comparisons/funke}: This folder contains four dynamic texture synthesis comparisons between the two-stream model and a recent (unpublished) approach \cite{funke2017}. The dynamic textures chosen are those reported by Funke \etal \cite{funke2017} which exhibit spatiotemporal homogeneity. For ease of comparison, the results from both models have been concatenated with their corresponding targets.
	\item \path{comparisons/xie_and_funke}: This folder contains nine dynamic texture synthesis comparisons between the two-stream model, Funke \etal's \cite{funke2017}, and Xie \etal's \cite{xie2017synthesizing}. The dynamic textures chosen cover the full range of the appearance and dynamics groupings listed in Sec.\ \ref{sec:user_study}. For ease of comparison, the results from all models have been concatenated with their corresponding targets.
\end{itemize}